Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-01 Paper/paper.tex
Narrowing down interesting parameter regions for robustness tests, adding figures that show Laplacian for my model of a perfect unweighted periodic adjacency matrix

Repetitive, periodic motions are ubiquitous in our world, such as animal locomotion (walking/flapping/slithering/swimming), mechanical motions (wheels spinning, pistons oscillating), biological rhythms (heartbeat, breathing), musical rhythms (drumming, feet tapping), etc.  In this work, we consider the problem of synthesizing a fine detail, ``slow motion'' template cycle of motion from a video of such a motion with many periods.  This fine detailed analysis can be applied, for example, to characterize subtle progressions of blood flow in the face during a heartbeat \cite{kumar2015distanceppg}.  Having access to a consensus template can also be used to visualize variations from cycle to cycle, such as in a repetitive automated action on an assembly line, where too much variation could indicate the onset of failure (TODO: citation), or in athletic activities for optimizing performance.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{BlockDiagram.pdf}
\caption{An block diagram of our technique, illustrated by a 1D periodic time series with additive Gaussian noise ($n_t \in \mathcal{N}(0, 1)$).  We estimate the period to be 24 (a multiple of the fundamental frequency), and perform a 24-length sliding window embedding.  From this, we use the mean of the birth and death time of the largest persistence dot to autotune the scale $\sigma$ for the graph Laplacian, from which we derive circular coordinates $\phi$, which can be used to reorder the sliding windows and use them to vote on a final slow motion template.  The result is shown in the bottom rightmost plot, superimposed with a ground truth period shows as an orange dotted line.}
\label{fig:ConceptFigure}
\end{figure*}

@@

\begin{equation}
SW_{d, \tau}[X(t)] = \left[ \begin{array}{c} X(t) \\ X(t + \tau) \\ \vdots \\ X(t + (d-1)\tau)  \end{array} \right] \in \mathbb{R}^{3WHd}
\end{equation}

where $\tau$ is a step size and $d$ is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  As shown by Takens \cite{takens1981detecting}, a sliding window of dimension $d = 2m+1$ of even a single generic observation function (time series) of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(d-1) \tau$, or simply $d-1$ in our case, maximizes the roundness of the embeddings when $d-1 = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D ``surrogate signal,'' and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate $d$.  This is quite similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

\subsection{Laplacian Eigenmaps And Persistent Homology}
\label{sec:laplacian}

@@

For a circle graph, where the $A_{ij} = 1$ if $|i-j| = 1$, $i=1, j=N$, and $j=1, i=N$, or zero otherwise, the eigenvectors spanning the image of the graph laplacian are

\begin{equation}
v_{2k}[n] = \cos(2 \pi k n / N), v_{2k-1}[n] = \sin(2 \pi k n / N), k \geq 1
\end{equation}

%In other words, the graph Laplacian of an $N$-point circle graph recovers the $N$-dimensional discrete Fourier transform.

with corresponding eigenvalues $\lambda_k = 2 - 2 \cos(2 \pi k / N)$ (and $\lambda_0 = 0$ with corresponding eigenvector $v_0[n] = 1$).  When computing the eigenvectors of a circle graph numerically, therefore, $v_1$ and $v_2$ should be a sine and cosine, or orthogonal linear combinations therein.

TODO: Talk about circulant graphs instead

@@

@@

\subsection{Sharpening Postprocessing}

The results of the median reordering can be blurry

@@

\subsection{Qualitative Results}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{XTSliceJumpingJacks.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of two men doing jumping jacks and for reordered videos with and without median consensus.}
\label{fig:XTSliceJumpingJacks}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{XTSliceFan.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of a fan with only 6 frames per period and the corresponding median consensus reordered video.}
\label{fig:XTSliceFan}
\end{figure}

>>>>

Repetitive, periodic motions are ubiquitous in our world, such as animal locomotion (walking/flapping/slithering/swimming), mechanical motions (wheels spinning, pistons oscillating), biological rhythms (heartbeat, breathing), musical rhythms (drumming, feet tapping), etc.  In this work, we consider the problem of synthesizing a fine detail, ``slow motion'' template cycle of motion from a video of such a motion with many periods.  This fine detailed analysis can be applied, for example, to characterize subtle progressions of blood flow in the face during a heartbeat \cite{kumar2015distanceppg}.  Having access to a consensus template can also be used to visualize variations from cycle to cycle, such as in a repetitive automated action on an assembly line, where too much variation could indicate the onset of failure, in repetitive human motions to assess the stress of those motions \cite{greene2017visualizing}, and in athletic activities for optimizing performance.

@@

\begin{figure*}[h!]
\centering
\includegraphics[width=\textwidth]{BlockDiagram.pdf}
\caption{An block diagram of our technique, illustrated by a 1D periodic time series with additive Gaussian noise ($n_t \in \mathcal{N}(0, 1)$).  We estimate the period to be 24 (a multiple of the fundamental frequency), and perform a 24-length sliding window embedding.  From this, we use the mean of the birth and death time of the largest persistence dot to autotune the scale $\sigma$ for the graph Laplacian, from which we derive circular coordinates $\phi$, which can be used to reorder the sliding windows and use them to vote on a final slow motion template.  The result is shown in the bottom rightmost plot, superimposed with a ground truth period shown as an orange dotted line.}
\label{fig:ConceptFigure}
\end{figure*}

@@

\begin{equation}
SW_{\text{dim}, \tau}[X(t)] = \left[ \begin{array}{c} X(t) \\ X(t + \tau) \\ \vdots \\ X(t + (\text{dim}-1)\tau)  \end{array} \right] \in \mathbb{R}^{3WH(\text{dim})}
\end{equation}

where $\tau$ is a step size and ``dim'' is the dimension of the embedding.  In this work, we fix $\tau = 1$ for simplicity and drop $\tau$ in the subscript, and we also sample times $t$ only at integer frame indices.  As shown by Takens \cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function (time series) of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(\text{dim}-1) \tau$, or simply $\text{dim}-1$ in our case, maximizes the roundness of the embeddings when $\text{dim}-1 = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D ``surrogate signal,'' and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate dim.  This is quite similar to the diffusion maps + autocorrelation approach in \cite{tralie2017quasi} for analyzing periodic videos.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

\subsection{Laplacian Eigenmaps}
\label{sec:laplacian}

\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{CirculantExample.pdf}
\caption{An example of the unweighted graph Laplacian on an ideal adjacency matrix corresponding to 6 periods of length 14 (top row), and the adjacency matrix from a sliding window embedding of two men doing jumping jacks (bottom row).  The two eigenvectors corresponding the smallest nonzero eigenvalues of the graph Laplacian span sinusoids that are 90 degrees out of phase, and which reconstruct a circle when plotted against each other.}
\label{fig:CirculantExample}
\end{figure}

@@

Now suppose we have performed a sliding window $SW_{\text{dim}}[X[n]]$ for a video $X$, where $X[n] = X[n+kT]$ for some integers $l, k$ so that there are $kT$ points in the sliding window (that is, $X$ repeats itself exactly every $T$ frames $k$ times total), and that $d_{ij} = ||SW_{\text{dim}}[X[i]] - SW_{\text{dim}}[X[n]]||_2$.  In this case, for any $\epsilon > 0$ and $|i - j| = kT$ for some integer $k$, $A_{ij} = 1$.

@@

In practice, look for the pair of eigenvectors within the first 10 with smallest zerocrossings

\subsection{Persistent Homology}

@@

@@

\subsection{Qualitative Results}
\begin{figure}
\centering
\includegraphics[width=\columnwidth]{XTSliceJumpingJacks.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of two men doing jumping jacks and for reordered videos with and without median consensus.}
\label{fig:XTSliceJumpingJacks}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{XTSliceFan.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of a fan with only 6 frames per period and the corresponding median consensus reordered video.}
\label{fig:XTSliceFan}
\end{figure}
